Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies

### 2010-04-23 papers/wmt10/joshua-progress.tex
dyer's edits
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1623 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\section{Lattice Parsing}
\todo{Chris Dyer}

@@

@@

Statistical models in machine translation exhibit spurious ambiguity.
That is, the probability of an output string is split
among many distinct derivations (e.g., trees or
segmentations). In principle, the goodness of a
string is measured by the total probability of its
many derivations. However, finding the best string
(e.g., during decoding) is then NP-hard.
The first version of \joshua implemented the
Viterbi approximation that measures the goodness 
of a string using only its most probable derivation.

The Viterbi approximation is efficient, but ignores most of the derivations in the hypergraph.
We implemented the variational decoding \cite{variational-decoding-acl09}, which works as follows.
Given a foreign string, the MT system first produces a hypergraph, 
which encodes a probability distribution $p$ over 
possible output strings and their derivations.
The variational decoding constructs a second distribution $q \in \Q$ 
that approximates $p$ as well as possible, 
and then finds the best string according to $q$ (instead of $p$).
The $q$ distribution is parameterized by an $n$-gram model, under which 
the last two steps can be performed efficiently via dynamic programming.
In this way, variational decoding considers all the derivations but still
allows tractable decoding.

@@

We have also implemented the average Perceptron algorithm and the forest-reranking \cite{zhifei-forest-reranking-galebook}.
Since the reference translation may not be in the hypergraph due to pruning or inherent
defficiency of the translation grammar, we need to use an {\em oracle translation} (i.e., the translation in
the hypergraph that is most simmilar to the reference translation) as a surrogate for training.
We implemented the {\em oracle extraction} algorithm described by \newcite{oracle-extraction-naacl09}
for this purpose.

>>>>

@@

\section{Lattice Parsing}

The bottom-up parsing algorithm used to generate the translation hypergraph has been generalized to support translation from not only an input sentence but from a \emph{word lattice} with runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.  This enables the exploration of a distribution over input sentences during decoding, and the best translation is selected from among any of them.  This functionality is useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters.

@@

Statistical models in machine translation exhibit spurious ambiguity.
That is, the probability of an output string is split
among many distinct derivations (e.g., trees or
segmentations). In principle, the goodness of a
string is measured by the total probability of its
many derivations. However, finding the best string
during decoding is then NP-hard.
The first version of \joshua implemented the
Viterbi approximation that measures the goodness 
of a string using only its most probable derivation.

The Viterbi approximation is efficient, but it ignores most of the derivations in the hypergraph.
We implemented variational decoding \cite{variational-decoding-acl09}, which works as follows.
First, given a foreign string (or lattice), the MT system produces a hypergraph, 
which encodes a probability distribution $p$ over 
possible output strings and their derivations.
Second, a distribution $q$ is selected that approximates $p$ as well as possible but comes from a family of distributions $\mathcal{Q}$ where inference is tractable.  Third, the best string according to $q$ (instead of $p$) is found.
In our implementation, the $q$ distribution is parameterized by an $n$-gram model, under which 
the second and third steps can be performed efficiently and exactly via dynamic programming.
In this way, variational decoding considers all derivations in the hypergraph but still
allows tractable decoding.

@@

We have also implemented the average Perceptron algorithm and forest-reranking \cite{zhifei-forest-reranking-galebook}.
Since the reference translation may not be in the hypergraph due to pruning or inherent
defficiency of the translation grammar, we need to use an {\em oracle translation} (i.e., the translation in
the hypergraph that is most simmilar to the reference translation) as a surrogate for training.
We implemented the {\em oracle extraction} algorithm described by \newcite{oracle-extraction-naacl09}
for this purpose.
